\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-12 syntax.tex
Save.
Save. Example is added, please check.

Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{\cL\tau[\pi_1]}$.  Though we will still introduce rules for sealing judgement, and show that all these rules are derivable by term judgements.

@@

We have shown case some exemplar rules for all four standard type judgments, now we focus on the newly introduced facility.

@@

Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type. To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}. Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type.

@@

\judgebox{\goodSig{\Gamma}{\sigma}{n} }
$$
\Rule[name=Linkage Type]
{\goodSig{\Gamma}{\sigma}{n}}
{\goodType{\Gamma}{\cL \sigma}{i}}
\quad
\Rule[name=Sig/Linkage Subst]
{\goodSig{\Delta}{\sigma}{n}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodSig{\Gamma}{\sigma[\gamma]}{n}
  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}}
$$

$$
\Rule[name=Empty Sig]
{}
{\goodSig{\Gamma}{\nu\cdot}{0}}
\quad
\Rule[name=Sig Add]
{\goodSig{\Gamma}{\sigma}{n} 
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}
 \quad \goodType{\Gamma, A}{T}{i}}
{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
$$
$$
\Rule[name=Empty Lnkg]
{}
{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
\quad
\Rule[name=Lnkg Add]
{ \goodTerm{\Gamma}{o}{\cL \sigma} 
\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
 \quad \goodTerm{\Gamma, A}{t}{T}
}
{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
$$

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields. We omit the naming/label information for fields in the meta-theory. We can use $\cL$ to retrieve the corresponding types of linkage of a given signature. Our signature corresponds to Coq's Module Type and our linkage corresponds to Coq's Module.

The signature for linkage closely follows the signature for inductive type.  They are both constructed from the empty stuff, and respect substitution lemma for sure. We have projection for each componenet due to harmony \citep{pfenning2009lecture} (local soundness and completeness), but we omit them and the substitution laws for either concatenation and projection. For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.

Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the design of our linkage, so an obvious question is why is that sealing judgement there?

@@

\begin{minted}[]{agda}
Seal : (Γ : Con) → (σ : Sig Γ n) → (τ : Sig Γ n) → Set 
Seal Γ σ τ = Tm (Γ, 𝓛σ) 𝓛τ[p₁]

-[-]S : (f : Seal Γ σ₁ σ₂) → (τ : Tms Θ Γ) → Seal Θ σ₁[τ] σ₂[τ]
f[τ]S = t[τ ↑ 𝓛σ₁]

idₛ : Seal Γ σ σ
idₛ = π₁
\end{minted}

@@

Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e. we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption when needed.

>>>>

Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is the most special one because it is actually just a short hand of term judgement $\goodTerm{\Gamma , \cL \sigma}{\_}{A[\pi_1]}$.

@@

We have shown case some exemplar rules for all four standard type judgments, and now we focus on the newly introduced facility.

@@

Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type. To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}. Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig Proj} to extract and corresponding $A_i, B_i$.

@@

\judgebox{\goodSig{\Gamma}{\sigma}{n} }
$$
\Rule[name=Lnkg Type/Compile]
{\goodSig{\Gamma}{\sigma}{n}}
{\goodType{\Gamma}{\cL \sigma}{i}
\quad \goodType{\Gamma}{\cC \sigma}{}}
\quad
\Rule[name=Sig/Lnkg Subst]
{\goodSig{\Delta}{\sigma}{n}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodSig{\Gamma}{\sigma[\gamma]}{n}
  \quad \goodType{\Gamma}{\cL (\sigma[\gamma]) \equiv (\cL \sigma)[\gamma]}{i}}
$$

$$
\Rule[name=Ept Sig]
{}
{\goodSig{\Gamma}{\nu\cdot}{0}}
\quad
\Rule[name=Sig Add]
{\goodSig{\Gamma}{\sigma}{n} 
 \quad \goodSeal{\Gamma}{f}{\sigma}{A}
 \quad \goodType{\Gamma, A}{T}{i}}
{\goodSig{\Gamma}{(\nu^+ \ \sigma \ \{f\} \ T)}{n+1}}
$$

$$ 
\Rule[name=Sig Proj]
{\goodSig{\Gamma}{\sigma}{n+1}}
{\goodSig{\Gamma}{p_1\nu \ \sigma}{n}
\\ \goodType{\Gamma}{p_1\nu' \ \sigma}{}
\\ \goodSeal{\Gamma}{p_f\nu \  \sigma}{p_1\nu \  \sigma}{p_1\nu' \sigma}
\\ \goodType{\Gamma, p_1\nu' \ \sigma}{p_2\nu \ \sigma}{}
}
$$

$$
\Rule[name=Ept Lnkg]
{}
{\goodTerm{\Gamma}{\mu\cdot}{\cL \nu\cdot}}
\quad
\Rule[name=Lnkg Add]
{ \goodTerm{\Gamma}{o}{\cL \sigma} 
\quad  \goodSeal{\Gamma}{f}{\sigma}{A} 
 \quad \goodTerm{\Gamma, A}{t}{T}
}
{\goodTerm{\Gamma}{(\mu^+ \ o \ \{f\} \ t)}{\cL(\nu^+ \ \sigma \ \{f\} \ T)}}
$$

$$
\Rule[name=Lnkg Proj]
{\goodTerm{\Gamma}{o}{\cL\sigma}}
{\goodTerm{\Gamma}{p_1\mu \ o}{\cL (p_1\nu \ \sigma)}
\\ \goodTerm{\Gamma, p_1\nu' \ \sigma}{p_2\mu \ o}{p_2\nu \ \sigma}
}
\quad 
\Rule[name=Lnkg Compile]
{ \goodTerm{\Gamma}{o}{\cL \sigma} 
}
{
  \goodTerm{\Gamma}{\cCt o}{\cC \sigma}
}
$$

$$
\Rule[name=Compile]
{}
{\goodType{\Gamma}{\cC \nu\cdot \equiv \top}{} 
\\ \goodTerm{\Gamma}{ \cCt \mu\cdot \equiv ()}{\cL \nu\cdot}
\\
\goodType{\Gamma}{\cC (\nu^+ \ \sigma \ \{f\} \ T) \equiv 
    \Sigma (\cC \sigma) T[sf \ f]}{}
\\\\ \goodTerm{\Gamma}{\cCt \ (\mu^+ \ o \ \{f\}\ t) \equiv ((\cCt \ o), t[sf \ f][(id, \cCt \ o)]) }{}
}
$$
% $$
% \Rule
% { \goodSeal{\Gamma}{f}{\sigma}{A}
% \\ \goodType{\Gamma, A}{T}{}
% \\ \goodTerm{\Gamma}{t}{T}
% }{}
% $$

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature for linkages in the context $\Gamma$ with $n$ fields. We omit the naming/label information for fields in the meta-theory. We can use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).

The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.  They are both constructed from the empty stuff, and respect substitution lemma for sure. We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony \citep{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules). We omit them and the substitution laws for either field addition and projection. For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.

Intuitively, both of the $\mu^+, \nu^+$ rules should be motivated by the implementation of our linkage, so an obvious question is what is this sealing judgement for?

@@

\begin{minted}[]{agda}
Seal : (Γ : Con) → (σ : Sig Γ n) → (A : Ty Γ) → Set 
Seal Γ A τ = Tm (Γ, 𝓛σ) A[p₁]

-[-]S : (f : Seal Γ σ A) → (τ : Tms Θ Γ) → Seal Θ σ[τ] A[τ]
f[τ]S = t[τ ↑ A]

idₛ : Seal Γ σ 𝓒σ
idₛ = π₁
\end{minted}

@@

With the definition of sealing judgement in mind, we can see \ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the implementation. In the implementation, we will have "self__" all around when defining new fields in our plugin, and this "self__" is exactly the $A$ in the context of the two fields -- especially when we don't do any sealing, this $A = \cC \sigma$, which is exactly how the implementation expose "self__" as a module parameter.

Next we talk about inheritance judgement. Note that, there will be a lot of $\sigma_1, A_1$ and $\sigma_2, A_2$, and in these cases, $A_1$ is the result of the sealing of $\sigma_1$, i.e. we assume we have a term $\goodSeal{\Gamma}{f_1}{\sigma_1}{A_1}, \goodSeal{\Gamma}{f_2}{\sigma_2}{A_2}$ in the assumption of judgements when needed.
